---
title: "Kericho"
author: greyhypotheses
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: brown
citecolor: green
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
---



<!--- Global Settings --->
```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = 'latex')
options(knitr.kable.NA = '')
```


<!--- Libraries --->
```{r include = FALSE}
library(xtable)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(matrixStats)
library(lubridate)

library(geoR)
library(PrevMap)

library(magrittr)
library(latex2exp)
library(patchwork)
library(kableExtra)
```

The set of external functions used thus far - relative to, therefore based in, GitHub repository [premodelling/time](https://github.com/premodelling/time/tree/develop) - are

<!--- External Functions --->
\footnotesize
```{r echo = TRUE, warning = FALSE, message = FALSE}
  sys.source(file = 'R/kericho/StudyData.R', envir = knitr::knit_global())

  sys.source(file = 'R/kericho/problems/explore/ExplorationGraph.R',
             envir = knitr::knit_global())
  sys.source(file = 'R/kericho/problems/rainfall/Graphs.R',
           envir = knitr::knit_global())
  sys.source(file = 'R/kericho/problems/specified/PredictionsGraph.R',
             envir = knitr::knit_global())

  sys.source(file = 'R/functions/TimeDependentLag.R', envir = knitr::knit_global())
  sys.source(file = 'R/functions/ConfidenceInterval.R', envir = knitr::knit_global())

  sys.source(file = 'docs/programme/mathematics/auxiliary_function.R',
             envir = knitr::knit_global())
```
\normalsize

\vspace{35pt}

**Data Set-up**

The original data set, with appended time dependent variables, is

\vspace{10pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
instances <- StudyData()
str(object = instances)
```
\normalsize

\clearpage

# Exploring Relationships

An exploration of the relationship between ln(cases) and maximum temperature, minimum temperature, and rain.

```{r out.width = '60%', fig.height = 6.5, fig.align = 'center', fig.cap = 'the relationship between ln(cases) and maximum temperature, minimum temperature, and rain'}
ExplorationGraphNL(instances = instances)
```


\clearpage

# The Rainfall Series

The function *TimeDependentLag()* creates lagged fields.  Hence, expression

\vspace{10pt}

\footnotesize
```{r comment = '         ', echo = TRUE}
  dataset <- TimeDependentLag(
    frame = instances, frame.date = 'date', frame.date.granularity = 'month',
    variables = 'Rain', lags = seq(from = 0, to = 4) )
```
\normalsize

\vspace{10pt}

creates lagged rainfall series; appended to the original data set.

\vspace{10pt}

\footnotesize
```{r comment = '         '}
head(dataset$frame) %>% tibble()
```
\normalsize

\vspace{10pt}

The graphs of *fig. 2* illustrate the relationship between ln(cases) and each lagged rainfall series. The numeric suffix of each graph's title denotes the rain series lag, in months.

\vspace{10pt}

```{r out.width = '65%', fig.height = 5.0, fig.align = 'center', fig.cap = 'ln(cases) and the lagged rainfall series.  the lags range from 0 to 4 months.'}
RainSeriesGraphs(data = dataset$frame, lagfields = dataset$laggedfields)
```

\vspace{10pt}

The degree of correlation between ln(cases) and each lagged rainfall series is quantifiable via the Pearson Correlation Coefficient.  The correlation values are:

\vspace{10pt}

\footnotesize
```{r comment = '         ', digits = 3}
correlation <- function (variable) {
  y <- cor(x = dataset$frame$CasesLN, y = dataset$frame[variable], use = 'complete.obs', method = 'pearson') %>%
    data.frame()
}
correlations <- dplyr::bind_cols(lapply(X = dataset$laggedfields, FUN = correlation))
row.names(correlations) <- 'ln(cases)'
correlations
```
\normalsize


\clearpage

# Exploring a Specific Model

Considering the time series model

\begin{align}
Y(t) &= \beta_{0} + \beta_{1}t + \beta_{2}I(pmax({t - 50, 0})) + \beta_{3}I(t > 225) \nonumber \\
     & \qquad + \beta_{4} minT(t - k) + \beta_{5} maxT(t - k) + \beta_{6}Rain(t - k)  \nonumber \\
     & \qquad \qquad + \mathcal{W}(t) + Z(t)
\end{align}

for the Kericho malaria cases data, wherein

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{c p{0.45\linewidth}l}\footnotesize
variable & description \\ \hline
$t$ & time (months) \\
$minT$ & mininum temperature \\
$maxT$ & maximum temperature \\
$Rain$ & rainfall (millimetres) \\
$k$ & lag; $k = 2$ months \\
$\mathcal{W}(t)$ & A Matern process whereby $\kappa = 2.5$ \\
$Z(t)$ & Gaussian noise \\
\end{tabular}
\end{center}

\vspace{20pt}



The function *TimeDependentLag()* creates lagged fields.  Hence, the lagged minimum temperature, maximum temperature, and rain fields:

\vspace{10pt}

\footnotesize
```{r echo = TRUE, warning = FALSE, message = FALSE}

  variables <- c('minT', 'maxT', 'Rain')

  T <- TimeDependentLag(
    frame = instances, frame.date = 'date', frame.date.granularity = 'month',
    variables = variables, lags = seq(from = 2, to = 2))
  dataset <- T$frame
```
\normalsize

\vspace{20pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
str(object = dataset)
```
\normalsize


\clearpage

## Exercise 1: Model Fitting

\vspace{10pt}

Prior to fitting *Eq. 1*, records that have `NaN` values ...

\vspace{10pt}

\footnotesize
```{r echo = TRUE}
  condition <- !is.na(dataset$rain_lag_2) | !is.na(dataset$mint_lag_2) | !is.na(dataset$maxt_lag_2)
  excerpt <- dataset[condition, ]
```
\normalsize

\vspace{10pt}

Hence, via the `fit.matern()` function

\footnotesize
```{r  echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
fit2.5 <- fit.matern(
  form = as.formula(log(Cases) ~ time + I(pmax(time - 50, 0)) + I(time > 225)
    + mint_lag_2 + maxt_lag_2 + rain_lag_2),
  time = 'time',
  start.cov.pars = c(1,5),
  kappa = 2.5,
  data = excerpt,
  method = 'nlminb')
```
\normalsize

\vspace{10pt}

The summary of the model fitted for *Eq. 1* is

\vspace{10pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
estimates <- summary(fit2.5, log.cov.pars = TRUE)
estimates
```
\normalsize

\vspace{10pt}

The parameter $\sigma^{2}$, $\phi^{2}$, and $\tau^{2}$, and their confidence intervals, are:

\vspace{10pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
intervals <- ConfidenceInterval(estimates = estimates, log.cov.pars = TRUE)
intervals %>%
  round(digits = 3)
```
\normalsize

\clearpage

## Exercise 2: Predictions

\vspace{10pt}

The foci herein are the $ln(cases)$ point predictions, and their 95% prediction intervals, w.r.t. the months of the Kericho data set.  The $time.predict()$ function of *auxiliary_function.R*

\vspace{10pt}

\footnotesize
```{r comment = '         ', echo = TRUE}
predictor <- time.predict(
  fitted.model = fit2.5,
  predictors = excerpt[, c('time', 'mint_lag_2', 'maxt_lag_2', 'rain_lag_2')],
  time.pred = excerpt$time,
  scale.pred = 'exponential')
```
\normalsize

\vspace{10pt}

creates the $time.predict()$ object of predictions, including the confidence intervals. The resulting graph (*fig. 3*) illustrates the curve's confidence intervals band.

\vspace{10pt}

```{r out.width = '65%', fig.align = 'center', results = 'hide', fig.cap = 'Predictions: ln(cases) and confidence interval'}
PredictionsGraphNaturalLog(predictor = predictor, original = excerpt$CasesLN)
```

\vspace{10pt}



