---
title: "Kericho"
author: greyhypotheses
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: brown
citecolor: green
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
---



<!--- Global Settings --->
```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = 'latex')
options(knitr.kable.NA = '')
```


<!--- Libraries --->
```{r include = FALSE}
library(xtable)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(matrixStats)
library(lubridate)


library(magrittr)
library(latex2exp)
library(patchwork)
library(kableExtra)
```

The set of external functions used thus far - relative to, therefore based in, GitHub repository [premodelling/time](https://github.com/premodelling/time/tree/develop) - are

<!--- External Functions --->
\footnotesize
```{r echo = TRUE, warning = FALSE, message = FALSE}
  sys.source(file = 'R/kericho/StudyData.R', envir = knitr::knit_global())
  sys.source(file = 'R/kericho/functions/TimeDependentLag.R',
             envir = knitr::knit_global())
  sys.source(file = 'R/kericho/problems/fourth/PredictionsGraph.R',
             envir = knitr::knit_global())

  sys.source(file = 'docs/programme/mathematics/auxiliary_function.R',
             envir = knitr::knit_global())
```
\normalsize

\vspace{35pt}

# Problem 4

Considering the time series model

\begin{align}
Y(t) &= \beta_{0} + \beta_{1}t + \beta_{2}I(pmax({t - 50, 0})) + \beta_{3}I(t > 225) \nonumber \\
     & \qquad + \beta_{4} minT(t - k) + \beta_{5} maxT(t - k) + \beta_{6}Rain(t - k)  \nonumber \\
     & \qquad \qquad + \mathcal{W}(t) + Z(t)
\end{align}

for the Kericho malaria cases data, wherein

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{c p{0.45\linewidth}l}\footnotesize
variable & description \\ \hline
$t$ & time (months) \\
$minT$ & mininum temperature \\
$maxT$ & maximum temperature \\
$Rain$ & rainfall (millimetres) \\
$k$ & lag; $k = 2$ months \\
$\mathcal{W}(t)$ & A Matern processwhereby $\kappa = 2.5$ \\
$Z(t)$ & Gaussian noise \\
\end{tabular}
\end{center}

\vspace{20pt}

**Data Set-up**

The original data set, with appended time dependent variables, is

\vspace{10pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
  instances <- StudyData()
  str(object = instances)
```
\normalsize

\vspace{10pt}

The function *TimeDependentLag()* creates lagged fields.  Hence, the lagged minimum temperature, maximum temperature, and rain fields:

\vspace{10pt}

\footnotesize
```{r echo = TRUE, warning = FALSE, message = FALSE,}
  LaggedSeries <- function (variable) {
    temporary <- TimeDependentLag(frame = instances,
                                  frame.date = 'date',
                                  frame.date.granularity = 'month',
                                  frame.focus = variable,
                                  lags = seq(from = 2, to = 2) )
    series <- temporary$frame[temporary$lagfields]

    return(series)
  }
  lagged.variables <- lapply(X = c('minT', 'maxT', 'Rain'), FUN = LaggedSeries)
  lagged.variables <- dplyr::bind_cols(lagged.variables)
  instances <- cbind(instances, lagged.variables)
```
\normalsize

\vspace{20pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
str(object = instances)
```
\normalsize


\clearpage

## Exercise 1: Model Fitting

Prior to fitting *Eq. 1*, records that have `NaN` values ...

\vspace{10pt}

\footnotesize
```{r echo = TRUE}
  condition <- !is.na(instances$rain_lag_2) | !is.na(instances$mint_lag_2) |
    !is.na(instances$maxt_lag_2)
  excerpt <- instances[condition, ]
```
\normalsize

\vspace{10pt}

Hence, via the `fit.matern()` function

\footnotesize
```{r  echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
  fit2.5 <- fit.matern(
    form = as.formula(log(Cases) ~ time + I(pmax(time - 50, 0)) + I(time > 225)
      + mint_lag_2 + maxt_lag_2 + rain_lag_2),
    time = 'time',
    start.cov.pars = c(1,5),
    kappa = 2.5,
    data = excerpt,
    method = 'nlminb')
```
\normalsize

\vspace{10pt}

The summary of the model fitted for *Eq. 1* is ...

\vspace{20pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
  estimates <- summary(fit2.5, log.cov.pars = TRUE)
  estimates
```
\normalsize

\vspace{20pt}

The natural logarithm scale values of $\sigma^{2}$, $\phi^{2}$, and $\tau^{2}$, and their confidence intervals:

\vspace{10pt}

\footnotesize
```{r echo = TRUE}
  parameters <- data.frame(estimates$cov.pars)
  parameters$interval <- qnorm(p = 0.975, lower.tail = TRUE) * parameters$StdErr
  parameters[, c('lower_ci', 'upper_ci')] <- parameters$Estimate +
    matrix(data = parameters$interval) %*%  matrix(data = c(-1, 1), nrow = 1, ncol = 2)
```
\normalsize

\vspace{10pt}

Consequently, their normal scale values are

\vspace{10pt}

\footnotesize
```{r echo = TRUE}
  parameters[, 'exp(Estimate)'] <- exp(parameters$Estimate)
  parameters[, c('exp(lower_ci)', 'exp(upper_ci)')] <-
    as.matrix(exp(parameters[, c('lower_ci', 'upper_ci')]))
```
\normalsize

\vspace{10pt}

Hence

\vspace{10pt}

\footnotesize
```{r comment = '         ', warning = FALSE, message = FALSE}
  parameters %>%
          dplyr::select(!interval) %>%
          round(digits = 3)
```
\normalsize

\clearpage

## Exercise 2: Predictions

The foci herein are the $ln(cases)$ point predictions, and their 95% prediction intervals, w.r.t. the months of the Kericho data set.  Using the $time.predict()$ function of *auxiliary_function.R*:

\vspace{10pt}

\footnotesize
```{r comment = '         ', echo = TRUE}
  predictor <- time.predict(
    fitted.model = fit2.5,
    predictors = excerpt[, c('time', 'mint_lag_2', 'maxt_lag_2', 'rain_lag_2')],
    time.pred = excerpt$time,
    scale.pred = 'exponential')
```
\normalsize

\vspace{10pt}

creates the $time.predict()$ object of predictions, including the confidence intervals. The resulting graph ...

\vspace{10pt}

```{r out.width = '65%', fig.align = 'center', results = 'hide', fig.cap = 'Predictions: ln(cases) and confidence interval'}
PredictionsGraph(predictor = predictor)
```

\vspace{10pt}

